Effects of prenatal CORT and temperature on the response to a stressor

### 2025-06-09 ms.qmd
updates

---
docx: 
reference-doc: "./bib/tmpl.docx"
bibliography: "./bib/refs.bib"
execute:
  echo: false
  error: false
  cache: false
  warning: false
crossref:  
  fig-title: Figure    # (default is "Figure")
  title-delim: —     # (default is ":")
  fig-prefix: Fig.   # (default is "Figure")
  tbl-prefix: Tab.   # (default is "Table")
editor_options: 
  chunk_output_type: console
---

@@

We used two species of skinks, the common garden skink (*Lampropholis guichenoti*) and the rainbow skink or delicate skink (*L. delicata*). These are oviparous, generalist, small (∼35--55 mm snout-vent length (SVL)) lizards that are sympatric in suburban areas throughout south-eastern Australia ([@chapple_know_2011]). Both species have similar breeding periods, but reproductive output in *L. delicata* is one clutch of 1 to 6 eggs per season, while *L. guichenoti* lays 1-5 eggs per clutch, and usually two clutches per season ([@chapple_know_2011]; [@chapple_biology_2014]). This difference in reproductive output through the breeding season may involve different degree of sensitivity to maternal effects and its interaction with other aspects of early environment.

@@

## Results
```{r, models_mass}
#| label: models_mass
#
# A) L. delicata
deli_mod_mass <- brm(rescaled_delta_mass ~ food_ingested+cort*temp, data = deli_stressor_2, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
  plot(deli_mod)
  summary(deli_mod)
# Delta mass L. guichenoti
  guich_mod <- brm(rescaled_delta_mass ~ food_ingested+cort*temp, data = guich_stressor_2, family = gaussian(link = "identity"), chains = 4, cores = 4, iter = 2000, warmup = 1000, control = list(adapt_delta = 0.99))
  summary(guich_mod)
bayes_R2(guich_mod)
```

@@

We calculated $R^2$ as follows

$$
R^2 = \frac{SS_{reg}}{SS_{tot}}
$$ {#eq-r2}

@eq-r2

```{r, fig1}
#| label: fig-fig1
#| fig-cap: "Activity budget of the focal group of chacma baboons (Papio ursinus) at the Cape Peninsula, South Africa, during the study period. The data are presented as the mean percentage of time spent in each activity category per hour (± SE)."

data %>% ggplot(aes(x = Lizard_id))  + geom_histogram()

We have a total of `r length(unique(data$Lizard_id))`

```{r,tb1}
#| label: tbl-tb1
#| tbl-cap: "Summary of the data"

tab <- data %>% 
  group_by(Lizard_id) %>% 
  summarise(
    n = n()
  ) 
flextable(tab[2:3,])
```

# References

>>>>

---
title: "Effects of prenatal CORT and temperature on the response to a stressor"
format:
  docx: 
    bibliography: "./bib/refs.bib"
    csl: "./bib/animal-behaviour.csl"
    reference-doc: "./bib/tmpl.docx"
execute:
  echo: false
  error: false
  cache: false
  warning: false
link-citations: true
crossref:  
  fig-title: Fig    # (default is "Figure")
  title-delim: —     # (default is ":")
  fig-prefix: Fig.   # (default is "Figure")
  tbl-prefix: Table  # (default is "Table")
editor_options: 
  chunk_output_type: console
---

@@

We used two species of skinks, the common garden skink (*Lampropholis guichenoti*) and the rainbow skink or delicate skink (*L. delicata*). These are oviparous, generalist, small (∼35--55 mm snout-vent length (SVL)) lizards that are sympatric in suburban areas throughout south-eastern Australia ([@chapple_know_2011]). Both species have similar breeding periods, but reproductive output in *L. delicata* is one clutch of 1 to 6 eggs per season, while *L. guichenoti* lays 1-5 eggs per clutch, and usually two clutches per season ([@chapple_know_2011]; [@chapple_biology_2014]). **This difference in reproductive output through the breeding season may involve different degree of sensitivity to maternal effects and its interaction with other aspects of early environment.**

@@

## Results

```{r, models_behaviour}
#| label: models_behaviour
source(here("R", "func.R"))
#
refit <- FALSE
#
# A) L. delicata
deli_beh <- behav_clean %>% 
  filter(species == "delicata") %>%
  mutate(lat_move = lat_move + 1, 
         lat_shelter = lat_shelter + 1)
#
deli_move <- fit_m(df = deli_beh, fam = lognormal(), type = "beh", sp = "deli", var = "lat_move")
deli_shelter <- fit_m(df = deli_beh, fam = lognormal(), type = "beh", sp = "deli", var = "lat_shelter")
deli_emergence <- fit_m(df = deli_beh, fam = bernoulli(link = "logit"), type = "beh", sp = "deli", var = "emergence")
#
#
# B) L. guichenoti
guich_beh <- behav_clean %>%
  filter(species == "guichenoti") %>%
  mutate(lat_move = lat_move + 1, 
         lat_shelter = lat_shelter + 1)
#
guich_move <- fit_m(df = guich_beh, fam = lognormal(), type = "beh", sp = "guich", var = "lat_move")
guich_shelter <- fit_m(df = guich_beh, fam = lognormal(), type = "beh", sp = "guich", var = "lat_shelter")
guich_emergence <- fit_m(df = guich_beh, fam = bernoulli(link = "logit"), type = "beh", sp = "guich", var = "emergence")
```

```{r, models_mass}
#| label: models_mass
source(here("R", "func.R"))
#
refit <- FALSE
#
# A) L. delicata
deli_mass <- mass_clean %>%
  filter(species == "delicata")
#
deli_mass <- fit_m(df = deli_mass, fam = lognormal(), type = "mass", sp = "deli", var = "delta_mass_rescaled")
# 
#
# B) L. guichenoti
guich_mass <- mass_clean %>%
  filter(species == "guichenoti")
#
guich_mass <- fit_m(df = guich_mass, fam = lognormal(), type = "mass", sp = "guich", var = "delta_mass_rescaled")
```

```{r, clean_posteriors}
#| label: clean_posteriors
source(here("R", "func.R"))
#
variables <- c("move", "shelter", "emergence", "mass")
species <- c("deli", "guich")
# Initialize empty lists
posterior_list <- list(deli = list(), guich = list())
# Loop over species and variables
for(i in species){
  for(j in variables){
    model_object <- get(paste0(i, "_", j))
    result <- tidy_post(df = model_object)
    posterior_list[[i]][[j]] <- result
  }
}
#
deli_post <- dplyr::bind_rows(posterior_list$deli, .id = "variable")
guich_post <- dplyr::bind_rows(posterior_list$guich, .id = "variable")
```

@@

## Discussion

## References

## Supplementary material

_Table **S1**. Estimates of each predictor for each variable modelled for L. delicata.  Lower and upper CI, and pMCMC values test the hypothesis that the estimate is different from 0. In bold, the values that are significant at pMCMC < 0.05._

```{r, tablesummary_deli}
#| label: tablesummary_deli
#
# Table defaults
set_flextable_defaults(
 font.family = "Times New Roman",
 font.size = 10)
# Make Table
table_summ_deli <- flextable(deli_post) %>%
  flextable::compose(i = 1, j = 1, value = as_paragraph("Variable"), part = "header") %>% # Change the first column header
  align(align = "center", j = c(3:6), part = "body") %>%
  align(align = "center", j = c(1:6), part = "header") %>%
  bold(~`pMCMC` < 0.05, j = c("pMCMC", "CI_upper", "CI_lower", "Predictor"),
       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
  bold(~`pMCMC` <0.001, j = c("pMCMC", "CI_upper", "CI_lower", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
  flextable::compose(i = c(2:10,12:20,22:30,32:40), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
  flextable::hline(i = c(10,20,30), part = "body") %>% # Add horizontal lines
  autofit()
#
table_summ_deli
```

```{r, results='asis', echo=FALSE}
cat("\\newpage")
```

_Table **S1**. Estimates of each predictor for each variable modelled for L. delicata.  Lower and upper CI, and pMCMC values test the hypothesis that the estimate is different from 0. In bold, the values that are significant at pMCMC < 0.05._

```{r, tablesummary_guich}
#| label: tablesummary_guich
#
# Table defaults
set_flextable_defaults(
 font.family = "Times New Roman",
 font.size = 10)
# Make Table
table_summ_guich <- flextable(guich_post) %>%
  flextable::compose(i = 1, j = 1, value = as_paragraph("Variable"), part = "header") %>% # Change the first column header
  align(align = "center", j = c(3:6), part = "body") %>%
  align(align = "center", j = c(1:6), part = "header") %>% 
  bold(~`pMCMC` < 0.05, j = c("pMCMC", "CI_upper", "CI_lower", "Predictor"),
       bold = TRUE) %>%  # Bold when PMCMC is "<0.05"
  bold(~`pMCMC` <0.001, j = c("pMCMC", "CI_upper", "CI_lower", "Predictor")) %>%  # Bold when PMCMC is "<0.001"
  flextable::compose(i = c(2:10,12:20,22:30,32:40), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
  flextable::hline(i = c(10,20,30), part = "body") %>% # Add horizontal lines
  autofit()
#
table_summ_guich
```
